---
title: "iseder_HW_3"
author: "Izzy Sederbaum"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  bookdown::html_document2:
    theme: cosmo
    toc: true
    toc_float: true
bibliography: MyCollection.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(captioner)
table_nums  <- captioner(prefix = "Table")
figure_nums <- captioner(prefix = "Figure")
library(pander)
knitr::write_bib(x = c("rmarkdown", "knitr"), file = "MyCollection.bib")
```

## trying to upload this to git! 

```{r load-data, cache=TRUE}
load("/Users/izzy/Downloads/Add_Health_Wave1_All.Rdata")

addhealth <- Add_Health_Wave1_All %>% 
  select(h1gi6a, h1gi6b, h1gh1, h1da6, h1da8)
```

```{r recode_data}
addhealth <- addhealth %>%  
  mutate(health = 
           as.factor(
             case_when(
               h1gh1 == 1 ~ "high",
               h1gh1 == 2 ~ "high",
               h1gh1 == 3 ~ "low",
               h1gh1 == 4 ~ "low",
               h1gh1 == 5 ~ "low"))) 

addhealth$health <- factor(addhealth$health, levels = c("low", "high"))


addhealth <- addhealth %>% 
  mutate(race = 
           as.factor(
             case_when(
               h1gi6a == 1 & h1gi6b == 0 ~ "White",
               h1gi6b == 1 & h1gi6a == 0 ~ "Black")))


addhealth <- addhealth %>% 
  mutate(exercise =
           as.factor(
             case_when(
               h1da6 == 0 ~ "no exercise this week",
               h1da6 == 1 ~ "exercised this week",
               h1da6 == 2 ~ "exercised this week",
               h1da6 == 3 ~ "exercised this week")))

addhealth <- addhealth %>% 
  filter(h1da8 < 996) %>% #filter out those who didn't answer
  mutate(TV = h1da8)

addhealth <- na.omit(addhealth)
```


## Descriptives

Here we're looking at race, exerice, tv watching and heatlh. `r table_nums(name = "healthtable", display = "cite")` and `r figure_nums(name = "addplot", display = "cite")` show the relationship between these variables, with `r table_nums(name = "healthtable", display = "cite")` showing the average amount of TV watched by race, self-rated health status and exercise.  

```{r table}
healthtable <- addhealth %>% 
  group_by(Race = race, 
           `Health Rating` = health,
           Exercise = exercise) %>% 
  summarise(TV = round(mean(TV), 2)) 

kable(healthtable, caption = "Table of Race, Heath Rating & TV Watching (in hours)") %>% 
  kable_styling(bootstrap_options = 
                      c("striped", "hover", "condensed", "responsive"), 
                  full_width = T, 
                  position = "center")
```


```{r,  fig.cap= "Boxplot of health rating and average TV watching by Race"}
addplot <- addhealth %>% 
  ggplot(aes(x=health, y= TV, fill=race)) +
  geom_boxplot() +
  xlab("Self Rated Health") +
  ylab("Hours of TV Watched per Week")

addplot
```



## Regression


Is there a relationship between exercise, TV watching, race and self-reported health? If I wanted to investigate, I would run the following regression:

$$health = \alpha + \beta_{1}\text{race} + \beta_{2}\text{exercise} + \beta_{3}\text{TV} + \varepsilon$$


Working in R Markdown has been a steep learning curve, but thankfully, sources such as [@rmarkdown2018] have been super helpful.


## Source Code

```{r}
cat(readLines("iseder_HW3.Rmd"), sep = '\n')
```

## References

---
nocite: '@*'
---
