---
title: "iseder_HW_10"
author: "Izzy Sederbaum"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  bookdown::html_document2:
    theme: cosmo
    toc: true
    toc_float: true
bibliography: worldbank.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(pander)
library(readxl)
library(kableExtra)
library(captioner)

```

```{r}
table_nums <- captioner(prefix = "Table")
figure_nums <- captioner(prefix = "Figure")
```

```{r import_data}
births <- read.table("C:/Users/iseder/OneDrive - UW/Year 2/Winter 2021/CSDE 502/Iceland.txt", header = T) 

births <- births %>%
  set_names(~ str_to_lower(.) %>%
              str_replace_all(" ", "_")) 

econ <- read_excel("C:/Users/iseder/OneDrive - UW/Year 2/Winter 2021/CSDE 502/Iceland_gdp.xls") 

```

```{r filter_years}
births <- births %>% 
  filter(year %in% c(1960:2018))

econ <- econ %>% 
  filter(year %in% c(1960:2018)) %>% 
  mutate(year = as.integer(year))
```

```{r bind_data}
iceland <- as.data.frame(left_join(births, econ))
```
```{r GDP_inbillions}
iceland <- iceland %>% 
  mutate(GDP = round(GDP/1000000000, digits = 2))
```
```{r decades}
iceland <- iceland %>% 
  mutate(decade = case_when(year %in% c(1960:1969) ~ "60s",
                            year %in% c(1970:1979) ~ "70s",
                            year %in% c(1980:1989) ~ "80s",
                            year %in% c(1990:1999) ~ "90s",
                            year %in% c(2000:2010) ~ "2000s",
                            year %in% c(2010:2018) ~ "2010s",
                            TRUE ~ "unknown") %>% 
  factor(ordered = TRUE))
```  
```{r reorder_decades}
iceland <- iceland %>% 
  mutate(decade = fct_relevel(decade, "60s", "70s", "80s", "90s", "2000s", "2010s"))
```
  
## Introduction

This paper looks GDP and number of births in Iceland, between 1960 and 2018, specifically looking at two possible relationships-- that between GDP and total births and GPD and the when the share of female births is greater than 50%. Birth data was dowloaded from [The Human Mortality Database](http://mortality.org) and GDP data was downloaded from [The World Bank](https://data.worldbank.org/indicator/NY.GDP.MKTP.CD?locations=IS). Iceland is a country located in the North Atlantic Ocean. It is the world's oldest democracy, but it is also ethnically homogeneous. Iceland's economy relies heavily on fishing and manufacturing, and about two-fifths of its GDP is exports [@Iceland].  

## Analysis

The purpose of this paper is to simply examine the possible relationship between number of births and GDP. As such, the descriptive data is displayed in the tables and graphs below. 

```{r}
table1 <- iceland %>%
  group_by(Decade = decade) %>% 
  summarise(`Total Girl Births` = sum(female),
            `Total Boy Births`  = sum(male),
            `Total Births`      = sum(total),
            `Average GDP (in Billions)` = round(mean(GDP), digits = 2)) 
  
kable(table1, caption = "Total Births & Average GDP in Each Decade, 1960-2018") %>% 
  kable_styling(full_width = T, position = "center") 
  
```

As we see in `r table_nums(name = "table1", display = "cite")` both the GDP of Iceland has been increasing since 1960, while the number of births has relatively stayed the same (it's worth noting that the data only go through 2018, which would explain the much lower number of births in that decade. When we look at birth by gender over the entire span of the dataset, (`r figure_nums(name = "figure1", display = "cite")`) we see that total births are relatively cyclical, which a high of `r max(iceland$total)` 2009. Previous highs occured in 1960, 2008, and 2010. We can also see that there are only 4 years where girls make up a larger share of births than boys. 

```{r, fig.cap= "Births by Gender: 1960-2018"}
figure1 <- ggplot(data = iceland,
 aes(x = year)) +
  geom_line(aes(y = female, color = "female"), size = 1) +
  geom_line(aes(y = male, color = "male"), size = 1) +
  ylab("Number of Births") +
  xlab("Year") +
  scale_color_manual("",
                     breaks = c("female", "male"),
                     values = c("goldenrod", "skyblue3")) +
  ggtitle("Births of Girls and Boys, 1960-2018") +
  theme(legend.position = "bottom")
 
figure1
```


```{r}
figure2 <- ggplot(data = iceland,
 aes(x = year)) +
  geom_line(aes(y = GDP), color = "maroon", size = 1) +
  ylab("GDP in Billions ($)") +
  xlab("Year") +
  geom_vline(xintercept = c(1964, 1985, 1999, 2013)) +
  ggtitle("Iceland GDP per Year in Billions($)", subtitle = "Years with More Girls than Boys Born Marked by Verticle Lines")
```


`r figure_nums(name= figure2, display = "cite")` below displays Iceland's GPD from 1960; years where girls were a greater share of births than boys are marked with vertical black lines. 


```{r} 
figure2 
```



```{r}
iceland <- iceland %>% 
  mutate(more_girls = if_else(female > male, "More Girls than Boys Born", "More Boys than Girls Born"))
```

From this graph we see that there does not appear to be anything particularly different about years in which girls made up the majority of births. 


```{r}
table2 <- iceland %>% 
  group_by(`Gender Share` = more_girls) %>% 
  summarize(`Average GDP (in Billions $)` = round(mean(GDP), digits = 2)) %>% 
  ungroup
```

(also shown in `r table_nums("table2", display = "cite")`). 
```{r}
kable(table2, caption = "Average GDP by Birth Gender Majority") %>% 
kable_styling(full_width = T, position = "center") 
```

## Discussion

While we might expect to see an inverse relationship between the number of births and GDP, it appears from the data above that for Iceland, that relationship may not exist. 


## Source Code

```{r}
cat(readLines("iseder_HW10.Rmd"), sep = '\n')
```

## References

---
nocite: '@*'
---
