---
title: "CSDE502 HW 9"
author: "Izzy Sederbaum iseder"
output: 
    bookdown::html_document2:
        number_sections: true
        self_contained: true
        code_folding: hide
        toc: true
        toc_float:
            collapsed: true
            smooth_scroll: false
    pdf_document:
        number_sections: true
        toc: true
        fig_cap: yes
        keep_tex: yes
urlcolor: blue 
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(captioner)

figure_nums <- captioner(prefix = "Figure")
table_nums <- captioner(prefix = "Table")

library(haven)
library(tidyverse)
library(readstata13)
library(kableExtra)
library(curl)
library(dplyr)
`%notin%` <- Negate(`%in%`)
```

___Explanation___:
This assignment is intended to give you more practice delving into the Add Health data set and in manipulating additional variables. 

___Instructions___: 

1. Make sure your Rmd file has no local file system dependencies (i.e., anyone should be able to recreate the output HTML using only the Rmd source file).
1. Make a copy of this Rmd file and add answers below each question. The code that generated the answers should be included, as well as the complete source code for the document.
1. Change the YAML header above to identify yourself and include contact information.
1. For any tables or figures, include captions and cross-references and any other document automation methods as necessary.
1. Make sure your output HTML file looks appealing to the reader.
1. Upload the final Rmd to your github repository.
1. Download [`assn_id.txt`](http://staff.washington.edu/phurvitz/csde502_winter_2021/assignments/assn_id.txt) and include the URL to your Rmd file on github.com.
1. Create a zip file from your copy of `assn_id.txt` and upload the zip file to the Canvas site for Assignment 9. ___The zip file should contain only the text file. Do not include any additional files in the zip file--everything should be able to run from the file you uploaded to github.com. Please use zip format and not 7z or any other compression/archive format.___


#
__Using the full household roster (you'll need to go back the full raw data source, [21600-0001-Data.dta](http://staff.washington.edu/phurvitz/csde502_winter_2021/data/21600-0001-Data.dta.zip)), create the following variables for each respondent. Document any decisions that you make regarding missing values, definitions, etc. in your narrative as well as in the R code.  Include a frequency tabulation and a histogram of each result.__

#read in data

```{r}
myUrl <- "http://staff.washington.edu/phurvitz/csde502_winter_2021/data/21600-0001-Data.dta.zip"
# zipfile in $temp
zipfile <- file.path(getwd(),"/data/", basename(myUrl))
# dta file in $temp
dtafile <- tools::file_path_sans_ext(zipfile)
# check if the dta file exists
if(!file.exists(dtafile)){
    # if the dta file doesn't exist, check for the zip file
    # check if the zip file exists, download if necessary
    if(!file.exists(zipfile)){
        curl::curl_download(url = myUrl, destfile = zipfile)
    }
    # unzip the downloaded zip file
    unzip(zipfile = zipfile, exdir = getwd())
}

# if the data set has not been read, read it in 
if(!exists("addhealth")){
   addhealth <- haven::read_dta(dtafile)
}

# lowercase column names
# colnames(adhealth) %>% str_to_lower()
```


#look through dataset for vars to keep 
```{r}

#dataset label
attributes(addhealth)$label
#variable label (this is visible in the viewer too)
attributes(addhealth$AID)$label

#look through dataset for variables to keep

ahcomplete_metadata <- bind_cols(
    varname = colnames(addhealth),
    varlabel = addhealth %>% map(~ attributes(.)$label) %>% unlist()
)

DT::datatable(ahcomplete_metadata)
```



#select vars to keep
```{r}

# create a subset of columns
dat <- addhealth %>%
    select(starts_with("H1HR3")) 

```

#For each household realtionship variable, code refused (96), legitmate skip (97), dont' know (98), and Not Applicable (99) to missing values

```{r}
dat$H1HR3A[dat$H1HR3A %in% c(96:99)]<-NA_real_
dat$H1HR3B[dat$H1HR3B %in% c(96:99)]<-NA_real_
dat$H1HR3C[dat$H1HR3C %in% c(96:99)]<-NA_real_
dat$H1HR3D[dat$H1HR3D %in% c(96:99)]<-NA_real_
dat$H1HR3E[dat$H1HR3E %in% c(96:99)]<-NA_real_
dat$H1HR3F[dat$H1HR3F %in% c(96:99)]<-NA_real_
dat$H1HR3G[dat$H1HR3G %in% c(96:99)]<-NA_real_
dat$H1HR3H[dat$H1HR3H %in% c(96:99)]<-NA_real_
dat$H1HR3I[dat$H1HR3I %in% c(96:99)]<-NA_real_
dat$H1HR3J[dat$H1HR3J %in% c(96:99)]<-NA_real_
dat$H1HR3K[dat$H1HR3K %in% c(96:99)]<-NA_real_
dat$H1HR3L[dat$H1HR3L %in% c(96:99)]<-NA_real_
dat$H1HR3M[dat$H1HR3M %in% c(96:99)]<-NA_real_
dat$H1HR3N[dat$H1HR3N %in% c(96:99)]<-NA_real_
dat$H1HR3O[dat$H1HR3O %in% c(96:99)]<-NA_real_
dat$H1HR3P[dat$H1HR3P %in% c(96:99)]<-NA_real_
dat$H1HR3Q[dat$H1HR3Q %in% c(96:99)]<-NA_real_
dat$H1HR3R[dat$H1HR3R %in% c(96:99)]<-NA_real_
dat$H1HR3S[dat$H1HR3S %in% c(96:99)]<-NA_real_
dat$H1HR3T[dat$H1HR3T %in% c(96:99)]<-NA_real_


#this code sums the true/false conditions. Doesnt seem to work with %in%
dat$numsisters=rowSums(select(dat,starts_with("H1HR3"))==8,na.rm=TRUE) 
dat$numbrothers=rowSums(select(dat,starts_with("H1HR3")) ==5,na.rm=TRUE) 
dat$hhmembers=rowSums(select(dat,starts_with("H1HR3")) >0,na.rm=TRUE) 
dat$numsiblings=dat$numbrothers+dat$numsisters

dat$nummother=rowSums(select(dat,starts_with("H1HR3"))==14,na.rm=TRUE) 
dat$numfather=rowSums(select(dat,starts_with("H1HR3"))==11,na.rm=TRUE) 

dat$numparents=dat$numfather+dat$nummother

#replace children who have more than two parents with two parents

dat$numparents[dat$numparents>2]<-2

dat$numhhmembers_other=dat$hhmembers-dat$numbrothers-dat$numsisters-dat$numparents

```


##
__Total number in household__
table(dat$hhmembers)


##
__Number of sisters__

table(dat$numsisters)


##
__Number of brothers__

table(dat$numbrothers)

##
__Total number of siblings__

table(dat$numsiblings)

#
__What proportion of students live with two biological parents? Include the analysis in your R code.__

dat %>% 
    group_by(numparents) %>% 
    summarise(n=n()) %>% 
    mutate(pct = (n / sum(n) * 100) %>% round(1)) %>% 
    kable(caption="Proportion of Students who live with 1 or 2 Biological Parents)

#
__Calculate the number of household members that are NOT biological mother, biological father, full brother or full sister. Create a contingency table and histogram for this variable.__
```{r}
dat %>% 
    group_by(hhmembers,numhhmembers_other) %>% 
    summarise(n=n()) %>% 
    spread(numhhmembers_other,n) %>% 
    kable(caption="Number of Household Members that are Not Biological Parent or Full Sibling by Number of Total Household Members")
    
    hist(dat$numhhmembers_other)
```



## Source code
```{r comment=''}
cat(readLines(con = "iseder_HW_9.Rmd"), sep = '\n')
```
